Instant Continuation Marks--Just Add $\lambda$

### 2012-10-12 paper.tex
Syntax highlighting and small corrections.

\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{slatex}
\usepackage{listings}

@@

To avoid this roadblock, we instead take a macro-style approach; that is, we enhance the core language with facilities to manipulate continuation marks and desugar the enhanced language back to the core. To make our desugaring transformation portable to other languages, we define it over the \lc, the common core of most higher-order languages. With such a transformation, language semanticists do not need to reconcile the feature with other features in the language (provided they have already done so with $\lambda$) and their compiler writers do not need to worry about complicating their implementations (for the same reason).

@@

With a reasonable intuition and definitions hardened by random testing, we prove correctness of the continuation-passing variant of the transform $\mathcal{C}_{\mathrm{cps}}$.

@@

A properly-recursive factorial can be implemented in this JavaScript extension as
\begin{verbatim}
var fac = function( n ) {
  if( n == 0 ) {
    console.log( ccm );
    return 1;
  }
  else {
    return wcm { n }
               { n * fac( n - 1 ) };
  }
}
\end{verbatim}

A direct-style transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
\begin{verbatim}
var fac = function( flag, marks ) {
  return function( n ) {
    if( n == 0 ) {
      console.log( (function( flag, marks ) { return marks; })( false, marks ) );
      return 1;
    }
    else {
      return (function( flag, marks ) {
        return (function( mark_value, rest_marks ) {
          return n * fac( false, [ mark_value ].concat( rest_marks ) )( n - 1 );
        })( n, flag ? marks.slice( 1 ) : marks );
      })( flag, marks );
    }
  }
}
\end{verbatim}

As expected, the output of a manual initiation of this function is
\begin{verbatim}
> fac( false, [] )( 5 )
[ 1, 2, 3, 4, 5 ]
120
\end{verbatim}

A tail-recursive factorial is expressible in this extension as
\begin{verbatim}
var fac = function( n, acc ) {
  if( n == 0 ) {
    console.log( ccm );
    return acc;
  }
  else {
    return wcm { n }
               { fac( n - 1, n * acc ) };
  }
}
\end{verbatim}
and transforms to
\begin{verbatim}
var fac = function( flag, marks ) {
  return function( n, acc ) {
    if( n == 0 ) {
      console.log( (function( flag, marks ) { return marks; })( false, marks ) );
      return acc;
    }
    else {
      return (function( flag, marks ) {
        return (function( mark_value, rest_marks ) {
          return fac( true, [ mark_value ].concat( rest_marks ) )( n - 1, n * acc );
        })( n, flag ? marks.slice( 1 ) : marks );
      })( flag, marks );
    }
  }
}
\end{verbatim}

The output of a manual initiation of this function is
\begin{verbatim}
> fac( false, [] )( 5, 1 )
[ 1 ]
120
\end{verbatim}

>>>>

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{slatex}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{xcolor}

\lstset{
	basicstyle=\small\ttfamily,
	keywordstyle=\color{RoyalBlue},
	identifierstyle=\color{gray}
}

@@

To avoid this roadblock, we instead take a macro-style approach; that is, we enhance a core language with facilities to manipulate continuation marks and desugar the enhanced language back to the core. To make our desugaring transformation portable to other languages, we define it over the \lc, the common core of most higher-order languages. With such a transformation, language semanticists do not need to reconcile the feature with other features in the language (provided they have already done so with $\lambda$) and their compiler writers do not need to worry about complicating their implementations (for the same reason).

@@

With a reasonable intuition and definitions hardened by random testing, we prove correctness of the continuation-passing variant of the transform, $\mathcal{C}_{\mathrm{cps}}$.

@@

A properly-recursive factorial can be implemented in this JavaScript extension as
\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
var fac = function( n ) {
  if( n == 0 ) {
    console.log( ccm );
    return 1;
  }
  else {
    return wcm { n }
               { n * fac( n - 1 ) };
  }
}
\end{lstlisting}

A direct-style transformation of this function desugars these constructs into vanilla JavaScript, leveraging native arrays for lists of continuation marks.
\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
var fac = function( flag, marks ) {
  return function( n ) {
    if( n == 0 ) {
      console.log( (function( flag, marks ) { return marks; })( false, marks ) );
      return 1;
    }
    else {
      return (function( flag, marks ) {
        return (function( mark_value, rest_marks ) {
          return n * fac( false, [ mark_value ].concat( rest_marks ) )( n - 1 );
        })( n, flag ? marks.slice( 1 ) : marks );
      })( flag, marks );
    }
  }
}
\end{lstlisting}

As expected, the output of a manual initiation of this function is
\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
> fac( false, [] )( 5 )
[ 1, 2, 3, 4, 5 ]
120
\end{lstlisting}

A tail-recursive factorial is expressible in this extension as
\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
var fac = function( n, acc ) {
  if( n == 0 ) {
    console.log( ccm );
    return acc;
  }
  else {
    return wcm { n }
               { fac( n - 1, n * acc ) };
  }
}
\end{lstlisting}
and transforms to
\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
var fac = function( flag, marks ) {
  return function( n, acc ) {
    if( n == 0 ) {
      console.log( (function( flag, marks ) { return marks; })( false, marks ) );
      return acc;
    }
    else {
      return (function( flag, marks ) {
        return (function( mark_value, rest_marks ) {
          return fac( true, [ mark_value ].concat( rest_marks ) )( n - 1, n * acc );
        })( n, flag ? marks.slice( 1 ) : marks );
      })( flag, marks );
    }
  }
}
\end{lstlisting}

The output of a manual initiation of this function is
\begin{lstlisting}[language=C,morekeywords={var,function,wcm,ccm,true,false}]
> fac( false, [] )( 5, 1 )
[ 1 ]
120
\end{lstlisting}
